MLH1 Results Outline

### 2020-04-23 doc/Results_v2.rmd
wrote the big loop for triangle plots

- update figure

- make table, save in figure folder, print head here

- remove cody blocks

@@

#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
                 .(category),
                 summarize,

@@

DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "late Z"

@@

```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}
DMC1.DF$stage

@@

DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

@@

anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
#

@@

ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )

@@

ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )

@@

ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
```

@@

#high vs low group (L)
ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value =  0.002145

#high vs low group (Z)
ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value = 0.6628

#dom vs musc (L) (2 subsp vs)
ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "L")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "L")] )
# p-value = 0.003016

#dom vs musc (Z)
ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Z")] )
#p-value = 0.09588

#KAZ vs Dom (Z) (only 1 L cell)
ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Z")], 
      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Z")] )

@@

```{r DMC1_MLH1.MAIN_TABLE, echo=FALSE, warning = FALSE}

@@

##RATIOS
DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z

@@

```{r DMC1.table, echo=FALSE}

@@

DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
                   category == "WSB male" |
                     category == "KAZ male" |
                  category == "MSM male" |
                    category == "PWD male")
#attach this mini DF to the DMC1 data
DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count

@@

DMC1table.main <- merge(DMC1.target.strains, DMC1table, by="category")

@@

DMC1table.main99 <-  DMC1table.main[, c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]

@@

kable(DMC1table.main99, row.names = F) %>%
  kable_styling() %>%
  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
  scroll_box(width = "800px", height = "200px")
```

@@

```{r with.cow, echo=FALSE, warning=FALSE, message=FALSE}
library(cowplot)

top.triangle <- function() {
  plot( PER_Foci_2 ~ PER_Foci_1, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
      axes = FALSE, ylab = 'Foci2', xlab = NA, main='Foci1',
      xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
  axis(side = 2, las = 1, pos=0)
  axis(side = 3, las = 1, pos=1)
  abline(coef = c(0,1))

@@

#plot_grid(top.triangle, bottom.triangle, ncol=1)

#plot_grid(top.triangle, bottom.triangle, align = "v", rel_widths = c(1.3, 1.3))

```{r another.triangle, echo=FALSE, fig.width = 5, fig.height=5}

@@

```{r Q1.IFD.triangle, echo=FALSE}

#make triangle for the IFD measures (try doing female top male bottom)
#Curated_BivData[Curated_BivData$hand.foci.count == 2,]

#remove grey background
#add diagona

#trying with base R

plot( PER_Foci_2 ~ PER_Foci_1, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
      axes = FALSE, ylab = 'Foci2', xlab = NA, main='Foci1',
      xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )

axis(side = 2, las = 1, pos=0)
axis(side = 3, las = 1, pos=1)
abline(coef = c(0,1))

#bottom triangle, Female
plot( PER_Foci_1 ~ PER_Foci_2, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
      axes = FALSE, xlab = NA, ylab = NA, xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
axis(side = 1, las = 1, pos=0)
axis(4, las = 1, pos=1) #flip label to right side
abline(coef = c(0,1))

#use cowplot or another way to paste these together

```{r Q1.show.IFD_female.male.comparison, echo=FALSE}

#compare female IFD patterns which have telomeric bias (more similar to males)

@@

P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "WSB"),
    aes(y = Foci2, x= Foci1, color=strain))+
  geom_jitter(alpha=.5)+
  ggtitle("Raw IFD WSB")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank() )

>>>>

@@

#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Early",],
                 .(category),
                 summarize,

@@

DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Late",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "late Z"

@@

```{r MAIN.DMC1.plot, echo=FALSE, warning=FALSE}

@@

DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(alpha=.5) +geom_jitter()+ggtitle("DMC1 averages in juvenile mice\nZygotene spermatogcytes")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

@@

anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)

@@

ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "Early")],
               DMC1.DF$total.count[(DMC1.DF$stage == "Late")]  )

@@

ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Early")],
               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Late")]  )

@@

ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Early")],
               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Late")]  )
```

@@

#high vs low group (L)
ttest.HighLow.L.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Early") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value =  0.002145

#high vs low group (Z)
ttest.HighLow.Z.pval <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Late") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )$p.value
# p-value = 0.6628

#dom vs musc (L) (2 subsp vs)
ttest.subsp.L <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Early")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Early")] )
# p-value = 0.003016

#dom vs musc (Z)
ttest.subsp.Z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], DMC1.DF$total.count[(DMC1.DF$subsp == "Musc") & (DMC1.DF$stage == "Late")] )
#p-value = 0.09588

#KAZ vs Dom (Z) (only 1 L cell)
ttest.KAZ.Dom.z <- t.test(DMC1.DF$total.count[(DMC1.DF$subsp == "Dom") & (DMC1.DF$stage == "Late")], 
      DMC1.DF$total.count[(DMC1.DF$strain == "KAZ") & (DMC1.DF$stage == "Late")] )

@@

```{r DMC1_more.calqs, echo=FALSE, warning = FALSE}

@@

@@

```{r DMC1.table, echo=FALSE, message=FALSE}
##RATIOS
DMC1table.main$MLH1.L_ratio <- DMC1table.main$mean.MLH1/ DMC1table.main$mean.count.L 
DMC1table.main$MLH1.Z_ratio <- DMC1table.main$mean.MLH1 / DMC1table.main$mean.count.Z

@@

DMC1.target.strains <- subset(mean.MLH1.obs, category == "G male" |
                   category == "WSB male" |
                     category == "KAZ male" |
                  category == "MSM male" |
                    category == "PWD male")

#attach this mini DF to the DMC1 data
DMC1.mean.obs.L$mean.count.L <- DMC1.mean.obs.L$mean.count
DMC1.mean.obs.Z$mean.count.Z <- DMC1.mean.obs.Z$mean.count

@@

DMC1table.main88 <- merge(DMC1.target.strains, DMC1table, by="category")

@@

DMC1table.main99 <-  DMC1table.main99[,c("category","mean.MLH1","ncells.x","mean.count.L", "MLH1.L_ratio", "ncells.y","mean.count.Z", "MLH1.Z_ratio"),  drop=FALSE]

@@

kable(DMC1table.main99, row.names = F, caption = "DMC1 measures") %>%
  kable_styling() %>%
  add_header_above(c(" " = 2, "Early Zygotene" = 3, "Late Zygotene" = 3)) %>%
  scroll_box(width = "800px", height = "200px")

@@

```{r with.cow, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
library(cowplot)

top.triangle <- function() {
  plot( PER_Foci_2 ~ PER_Foci_1, data = (Curated_BivData %>% filter(sex == "female") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ), 
      axes = FALSE, ylab = 'Foci2', xlab = NA, main='Foci1',
      xlim=c(0,1), ylim=c(0,1),  xaxt="n", yaxt="n" )
  axis(side = 2, las = 1, pos=0)
  axis(side = 3, las = 1, pos=1)
  abline(coef = c(0,1))
}

@@

```{r another.triangle, echo=FALSE, fig.width = 5, fig.height=5, eval=FALSE}

@@

```{r updated.triangl.plots, echo=FALSE, message=FALSE, warning=FALSE}
my.data <- data.frame(my.x=c(.2, .4, .6, .1), my.y=c(.3, .5, .7, .9))
x.dist <- .5
my.data.2 <- transform(my.data, my.y=my.y + x.dist)

#x_dist is a value for spacing the trianlges, remeber that the data values need
#to be adjusted as well
#the rest of the positions are slightly off

#make a loop to enter/push the merge triangles into a empty list

#Top is female
top.triangle2 <- function() {
  #top-y position is male Pos2
  plot(  (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
    #top-x position is female Pos1     
    (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,

       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2) ), 
       col =  "coral4",
       axes=FALSE, ylab='', xlab="", 
       main='', xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)
  mtext("KAZ IFD Distributions", 3, 5, font=2, cex=1.3, adj=.5)

  mtext("Female Position.2", 2, .75)
  mtext("Female Position.1", 3, 2)
  axis(side=2, las=1, pos=0)
  axis(side=3, las=1, pos=1)
  lines(0:1, 0:1)
}

#bottom is male
bottom.triangle2 <- function() {
    #bottom-y position is male Pos1
  points(  (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
    #bottom-x position is male Pos2     
    (Curated_BivData %>% filter(strain == "KAZ") %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,

         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == "KAZ") %>% filter(hand.foci.count == 2) ),

    col =  "coral4",
         xpd=TRUE)
  mtext("Male Position.2", 1, 1.5, at=mean(par()$usr[1:2])) #  + x.dist
  mtext("Male Position.1", 4, 3)#, padj=par()$usr[1]
  x.at <- axisTicks(par()$usr[1:2], 0) #+ x.dist
  axis(side=1, las=1, pos=0, at=x.at, 
       labels=F, xpd=TRUE)
  mtext(seq(0, 1, .2), 1, 0, at=x.at)
  axis(4, las=1, pos=1) # + x.dist
  lines(0:1, 0:1, xpd=TRUE)# + x.dist

png("~./MLH1repo/doc/figureFiles/KAZ.triangle.png", width=650, height=500)
op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
top.triangle2()
bottom.triangle2()
par(op)
dev.off()

```{r auto.loop.ifd, echo=FALSE}

#don't think I need this loop
for(i in 1:length(auto_curate_DF$strain )){
  auto_curate_DF$quant_status[i] <- ifelse( is.element(auto_curate_DF$strain[i], exclude_mice_list), "exclude", NA)
  }

#big loop for automating the IFD triangles
only.male.strains <- c("F1", "PERC", "CZECH", "AST", "TOM", "HMI", "SPRET", "CAROLI", "other")
auto_curate_DF <- Curated_BivData[ ! Curated_BivData$strain %in% only.male.strains, ]

df_strain_list <- split(auto_curate_DF, (auto_curate_DF$strain))
df_strain_list <- df_strain_list[ -c(4,10,11,12,13,14,15,16,17,18,19) ]
#make a new empty df. create a ddply thing with mean calculations (this is like a list for matching up later)

for(i in 1:(length(df_strain_list))){
  current.strain <- as.character(df_strain_list[[i]]$strain[1])
  #declare file
  png(paste0("~./MLH1repo/doc/figureFiles/test.triangle_", current.strain, ".png"),  width=650, height=500)

  #Top is female
top.triangle2 <- function() {
  #top-y position is male Pos2
  plot(  (Curated_BivData %>% filter(strain == current.strain  ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_2 ~ 
    #top-x position is female Pos1     
    (Curated_BivData %>% filter(strain == current.strain   ) %>% filter(sex == "female") %>% filter(hand.foci.count == 2) )$PER_Foci_1,

       data= (Curated_BivData %>% filter(sex == "female") %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) ), 
       col =  colors_of_strains[current.strain],
       axes=FALSE, ylab='', xlab="", 
       main= "IFD Distribution",  xlim=c(0, 1), ylim=c(0, 1), xaxt="n", yaxt="n", asp=1)

  mtext(current.strain, 3, 5, font=2, cex=1.3, adj=.5)

  mtext("Female Position.2", 2, .75)
  mtext("Female Position.1", 3, 2)
  axis(side=2, las=1, pos=0)
  axis(side=3, las=1, pos=1)
  lines(0:1, 0:1)
  }

bottom.triangle2 <- function() {
    #bottom-y position is male Pos1
  points(  (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_1 ~ 
    #bottom-x position is male Pos2     
    (Curated_BivData %>% filter(strain == current.strain) %>% filter(sex == "male") %>% filter(hand.foci.count == 2) )$PER_Foci_2,

         data=(Curated_BivData %>% filter(sex == "male") %>% filter(strain == current.strain) %>% filter(hand.foci.count == 2) ),

    col =  colors_of_strains[current.strain], #this might not be working right
         xpd=TRUE)
  mtext("Male Position.2", 1, 1.5, at=mean(par()$usr[1:2])) #  + x.dist
  mtext("Male Position.1", 4, 3)#, padj=par()$usr[1]
  x.at <- axisTicks(par()$usr[1:2], 0) #+ x.dist
  axis(side=1, las=1, pos=0, at=x.at, 
       labels=F, xpd=TRUE)
  mtext(seq(0, 1, .2), 1, 0, at=x.at)
  axis(4, las=1, pos=1) # + x.dist
  lines(0:1, 0:1, xpd=TRUE)# + x.dist

  op <- par(mar=c(3, 4, 8, 12) + 0.1, oma=c(2, 0, 0, 2))
  top.triangle2()
  bottom.triangle2() #not working
  par(op)
  dev.off()
  }

```{r Q1.IFD_load.triangles, echo=FALSE, message=FALSE, warning=FALSE}

library(png)

G.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/G.triangle.png")
WSB.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/WSB.triangle.png")
LEW.triangle.png <- readPNG("~./MLH1repo/doc/figureFiles/LEW.triangle.png")

#make empty DF for ggplot object
empty.DF <- data.frame(col1 = c(1,3), col2 = c(1,3) )

#make an empty ggplot thing, then background ontop
WSB.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(WSB.triangle.png) + theme(axis.title.x=element_blank(),
 axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )

G.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(G.triangle.png) + theme(axis.title.x=element_blank(),
 axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )

LEW.triangle.ggplot <- ggplot(empty.DF, aes(x = col1, y = col2)) +
  background_image(LEW.triangle.png) + theme(axis.title.x=element_blank(),
 axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),
                                  axis.text.y=element_blank(),axis.ticks.y=element_blank() )

library("cowplot")
fig1  <- ggdraw() +
  draw_plot(WSB.triangle.ggplot, x=0, y=0, width=.45, height=.45) +
  draw_plot(G.triangle.ggplot,   x=0, y=.4, width=.45, height=.45) +
  draw_plot(LEW.triangle.ggplot, x=0, y=.6, width=.45, height=.45)

fig1

#ggsave("~./WildMiceMadison/results/Figure1.png")
```

@@

P1.by.P2_plot_raw.WSB <- ggplot(Curated_BivData.22 %>% filter(strain == "LEW"),
    aes(y = Foci2, x= Foci1, color=strain))+
  geom_jitter(alpha=.5)+
  ggtitle("Raw IFD WSB")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none") + ylim(c(0,160))+xlim(c(0,160))+ theme(  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank() )
